Effect of Black Hole Spins on Christodoulou Memory in Binary Mergers and its implication on Observing Gravitational Wave Bursts With Memory in Pulsar Timing Array

### 2019-10-09 notes/Paper.tex
Updated plots in the paper

\begin{figure}
 	\includegraphics[width=3.0in]{../plots/MemoryPlot_nonSpining/q7.pdf}
 	\caption{The figure above shows Christodoulou memory as a function of time during BBH coalesce. This figure shows a comparison for different mass ratios. It can be seen that higher mass ration tends to decrease the Chistodoulou memory. The early inspiral part is computed using post-Newtonian expression and the later part during BBH merger is computer using NR data. The red dashed line shows result from PN expression}
 	\label{fig:differentmassratio}
\end{figure}
\begin{figure}
	\includegraphics[width=3.0in]{../plots/MemoryPlot_AlignedSpinSXSdata/0p99.pdf}
	\includegraphics[width=3.0in]{../plots/MemoryPlot_AntialignedSpinSXSdata/m0p94.pdf}
	\caption{The figure above shows Christodoulou memory as a function of time during BBH coalesce. The top figure shows a comparison for different values of dimensionless spin parameter $\chi$. These results are for BBH system with equal spin magnitude for individual black hole and spins aligned with orbital angular momentum. It can be seen that higher spin magnitude tend to increase the Christodoulou memory. The bottom figure shows the anti aligned case.  }
	\label{fig:differntchivalues}
\end{figure}
\section{Observing Memory from merger of Super-massive black holes  }
Nonlinear memory from the merger of super massive black hole binary can be potentially observed in pulsar timing measurement. Gravitation wave, when they pass through the space between earth and pulsar effect the time of arrival of the radio pulses at earth. The modulation in the observed frequency is given by
\begin{equation}
\frac{\Delta \nu}{\nu_{0}} = \frac{1}{2c}\int_{0}^{D}d\lambda \left(e^{i}e^{j}\frac{\partial h_{ij}}{\partial t}\right) \Bigg\vert_{path}
\end{equation}

@@

The growth in the nonlinear memory signal amplitude from super massive black hole occurs on the time scale of a few days so the effect on timing residuals can be seen if observed over a few weeks period of time. A method for observing Nonlinear memory signal from super massive black hole mergers using pulsar timing has been proposed in ref\cite{Pshirkov2010}. The authors have shown that the present day pulsar timing precision allows one to detect BWM due to  SMBH merger at distance upto 1 Gpc (for case of equal mass $10^{8}$ solar mass). The details regarding the calculation of the timing residuals can be seen in \cite{Pshirkov2010}, where they consider the memory event as a step function which occurs in the middle of an observation time of 10 years. The method described in \cite{Pshirkov2010} is general and can be applied to case where the memory jump is not at step function. In our analysis we explore how the nonlinear memory grows during two weeks interval to time for different ranges of mass and spin parameter values. We also calculate the timing residuals as done in \cite{Pshirkov2010} and explore how it changes as a function of total mass and spin of black holes as shown in Fig[\ref{fig:MemoryMassVsSpin}].
\par 
Future observations with increased precision would be able to look at the memory growth on even shorter time of few weeks. The memory growth on the time scale of two weeks does not look like a step function, instead it's shape depends upon the spins of the merging black holes. In Fig[\ref{fig:resedualGrowth}].we show the memory growth interval in two weeks interval of time with solid lines for different range of spin and total mass of the binary. The dashed line show the complete memory growth in growth in geometrical units. It can be seen the as the total mass of binary becomes larger the two weeks interval of time shrinks in geometrical units. We have chosen this two weeks interval of time such that the change in memory during this time interval is maximum. For the pulsar timing observation we compute the residuals during this two week interval of time following the method described in \cite{Pshirkov2010}. We show a figure similar to
\begin{widetext}
\begin{figure}
	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksGeoUnits.pdf}
	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksInDays10pow8MSun.pdf}
	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksInDays10pow10MSun.pdf}
	\caption{The top figure shows the memory growth during merger of a binary with time measured in geometric units with total spin magnitude as indicated above. The two weeks interval of time for a given mass in indicated above. The above figure show the the memory growth in days. The left figure show this growth for $10^{8} M_{\odot}$ and the right figure show the the memory growth in days for $10^{10} M_{\odot}$}
	\label{fig:resedualGrowth}
\end{figure} 
\end{widetext}

>>>>

\begin{figure}
 	\includegraphics[width=3.0in]{../plots/MemoryPlot_nonSpining/q7.pdf}
 	\caption{The figure above shows Christodoulou memory as a function of time during BBH coalesce. This figure shows a comparison for different mass ratios. It can be seen that higher mass ration tends to decrease the Chistodoulou memory. The early inspiral part is computed using post-Newtonian expression and the later part during BBH merger is computer using NR data. The red dashed line shows result from PN expression}
 	\label{fig:differentmassratio}
\end{figure}
\begin{figure}
	\includegraphics[width=3.0in]{../plots/MemoryPlot_AlignedSpinSXSdata/0p99.pdf}
	\includegraphics[width=3.0in]{../plots/MemoryPlot_AntialignedSpinSXSdata/m0p94.pdf}
	\caption{The figure above shows Christodoulou memory as a function of time during BBH coalesce. The top figure shows a comparison for different values of dimensionless spin parameter $\chi$. These results are for BBH system with equal spin magnitude for individual black hole and spins aligned with orbital angular momentum. It can be seen that higher spin magnitude tend to increase the Christodoulou memory. The bottom figure shows the anti aligned case.  }
	\label{fig:differntchivalues}
\end{figure}
\section{Observing Memory from merger of Super-massive black holes  }
Nonlinear memory from the merger of super massive black hole binary can be potentially observed in pulsar timing measurement. Gravitational waves, when they pass through the space between earth and pulsar effect the time of arrival of the radio pulses at earth. The modulation in the observed frequency is given by
\begin{equation}
\frac{\Delta \nu}{\nu_{0}} = \frac{1}{2c}\int_{0}^{D}d\lambda \left(e^{i}e^{j}\frac{\partial h_{ij}}{\partial t}\right) \Bigg\vert_{path}
\end{equation}

@@

The growth in the nonlinear memory signal amplitude from super massive black hole occurs on the time scale of a few days so the effect on timing residuals can be seen if observed over a few weeks period of time. A method for observing Nonlinear memory signal from super massive black hole mergers using pulsar timing has been proposed in ref\cite{Pshirkov2010}. The authors have shown that the present day pulsar timing precision allows one to detect BWM due to  SMBH merger at distance upto 1 Gpc (for case of equal mass $10^{8}$ solar mass). The details regarding the calculation of the timing residuals can be seen in \cite{Pshirkov2010}, where they consider the memory event as a step function which occurs in the middle of an observation time of 10 years. The method described in \cite{Pshirkov2010} is general and can be applied to case where the memory jump is not at step function. In our analysis we explore how the nonlinear memory grows during two weeks interval to time for different ranges of mass and spin parameter values. We also calculate the timing residuals as done in \cite{Pshirkov2010} and explore how it changes as a function of total mass and spin of black holes as shown in Fig[\ref{fig:MemoryMassVsSpin}].
\par 
Future observations with increased precision would be able to look at the memory growth on even shorter time of few weeks. The memory growth on the time scale of two weeks does not look like a step function, instead it's shape depends upon the spins of the merging black holes. In Fig[\ref{fig:resedualGrowth}].we show the memory growth interval in two weeks interval of time with solid lines for different range of spin and total mass of the binary. The dashed line show the complete memory growth in growth in geometrical units. It can be seen the as the total mass of binary becomes larger the two weeks interval of time shrinks in geometrical units. We have chosen this two weeks interval of time such that the change in memory during this time interval is maximum. For the pulsar timing observation we compute the residuals during this two week interval of time following the method described in \cite{Pshirkov2010}. We show a figure similar to
%%\begin{widetext}
\begin{figure}
	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksGeoUnits.pdf}
	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksInDays10pow8MSun.pdf}
	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksInDays10pow9MSun.pdf}
	\includegraphics[width=3.0in]{../plots/PlotfromMathematicaData/MemoryGrowth2weeksInDays10pow11MSun.pdf}
	\caption{The top figure shows the memory growth during merger of a binary with time measured in geometric units with total spin magnitude as indicated above. The two weeks interval of time for a given mass as indicated above. The bottom three figures show the the memory growth in days for Mass $10^{8}M_{\odot}$, $10^{9}M_{\odot}$ and $10^{11}M_{\odot}$ respectively.}
	\label{fig:resedualGrowth}
\end{figure} 
%\end{widetext}
